### 2021-07-11 src/paper/appendix_data.tex
Start on extending the model fit section.

The model is described by a large number of parameters that govern the number of
contacts a person has, the likelihood of becoming infected on each contact, the
likelihood of developing light or strong symptoms or even dying from the disease as well
as the duration each stage of the disease takes.

Most of these parameters can be calibrated from existing datasets or the medical
literature or calibrated from surveys and empirical datasets.

@@

>>>>

The model is described by a large number of parameters that govern the number of contacts
a person has, the reduction in contacts due to NPIs, the demand for rapid tests and PCR
tests, the likelihood of becoming infected on each contact, the likelihood of developing
light or strong symptoms or even dying from the disease as well as the duration each
stage of the disease takes.

Many of these parameters can be calibrated from the medical literature or calibrated from
surveys and other empirical datasets.

@@

\input{paper/data/rapid_test_demand}

\input{paper/data/share_known_cases}

### 2021-07-11 src/paper/data/medical_parameters.tex

For the time from symptom onset until need for intensive care we rely on data by
\cite{Stokes2020}) and \href{https://bit.ly/3yKXFyu}{the OpenABM-Project}.

For those who will require intensive care we follow \citet{Chen2020} who estimate the
time from symptom onset to ICU admission as $8.5 \pm 4$ days. This aligns well with
numbers reported for the time from first symptoms to hospitalization:
\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of four. We assume
that the time between symptom onset and ICU takes four, six, eight or ten days with equal
probabilities.

@@

We take the survival probabilities and time to death and time until recovery from
intensive care from the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}. They report
time until death to have a mean of 11.74 days and a standard deviation of 8.79 days. To
match this approximately we discretize that 41\% of individuals who will die from
Covid-19 do so after one day in intensive care, 22\% day after twelve days, 29\% after 20
days and 7\% after 32 days. Again, we rescale this for every age group among those that
will not survive. For survivors the \href{https://tinyurl.com/y5owhyts}{OpenABM Project}
reports a mean duration of 18.8 days until recovery and a standard deviation of 12.21
days. We discretize this such that of those who recover in intensive care, 22\% do so
after one day, 30\% after 15 days, 28\% after 25 days and 18\% after 45 days.

>>>>

For the time from symptom onset until need for intensive care we rely on data by
\cite{Stokes2020}) and \citep{Hinch2021}.

For those who will require intensive care we follow \citet{Chen2020} who estimate the
time from symptom onset to ICU admission as $8.5 \pm 4$ days. This aligns well with
numbers reported for the time from first symptoms to hospitalization:

\citet{Gaythorpe2020} report a mean of 5.76 with a standard deviation of four. We assume
that the time between symptom onset and ICU takes four, six, eight or ten days with
equal probabilities.

@@

We take the survival probabilities and time to death and time until recovery from
intensive care from the \citet{Hinch2021}. They report time until death to have a mean
of 11.74 days and a standard deviation of 8.79 days. To match this approximately we
discretize that 41\% of individuals who will die from Covid-19 do so after one day in
intensive care, 22\% day after twelve days, 29\% after 20 days and 7\% after 32 days.
Again, we rescale this for every age group among those that will not survive. For
survivors the OpenABM Project reports a mean duration of 18.8 days until recovery and a
standard deviation of 12.21 days. We discretize this such that of those who recover in
intensive care, 22\% do so after one day, 30\% after 15 days, 28\% after 25 days and
18\% after 45 days.

### 2021-07-11 src/paper/data/number_of_contacts.tex
Join single plots and make single plots smaller plus various small improvements.

We calibrate the parameters for the predicted numbers of contacts from contact diaries
of over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
\citep{Mossong2008}. Each contact diary contains all contacts an individual had
throughout one day, including information on the other person (such as age and gender)
and information on the contact. Importantly, for each contact individuals entered of
which type the contact (school, leisure, work etc.) was and how frequent the contact
with the other person is.

Simplifying the number of contacts, we arrive at the following distributions of the
numbers of contacts by contact type.

\begin{figure}
    \centering
    % other contacts
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
        \caption{Number of Non Recurrent Other Contacts}
        \label{n_contacts_other_non_recurrent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
        \caption{Number of Daily Recurrent Other Contacts}
        \label{n_contacts_other_daily_recurrent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
        \caption{Number of Weekly Recurrent Other Contacts}
        \label{n_contacts_other_weekly_recurrent}
    \end{subfigure}

@@

    % work contacts
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
        \caption{Number of Non Recurrent Work Contacts}
        \label{n_contacts_work_non_recurrent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
        \caption{Number of Daily Recurrent Work Contacts}
        \label{n_contacts_work_daily_recurrent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
        \caption{Number of Weekly Recurrent Work Contacts}
        \label{n_contacts_work_weekly_recurrent}
    \end{subfigure}
    \vskip3ex

    \caption{Number of Contacts of the Different Contact Types}
    \label{fig:n_contacts_other}
    \floatfoot{\noindent
        \textit{Note:} The upper row shows the distribution of the number of other
        contacts individuals report. Other contacts include all contacts that are not
        household members, school contacts or work contacts, for example leisure
        contacts. The planned number of contacts is reduced by policies, seasonality and
        individual responses to events such as receiving a positive rapid test to the
        number of actual contacts with transmission potential.
        % non recurrent
        In the model it is sampled every day which of the numbers of non recurrent
        contacts a person is planned to have. Note that the contact diaries include such
        high values that super spreading events are well possible in our model through
        non recurrent other models. We assume that individuals in households with
        children or teachers or retired individuals have additional non recurrent
        contacts during school vacations to cover things like family visits or travel
        during vacations. We estimate this to be on average 0.5 additional contacts per
        vacation day.
        % daily and weekly
        For the recurrent other contacts, individuals are assigned to groups that are
        time constant and that meet daily or weekly. The share of individuals who attend
        in a way that has transmission potential is reduced by policies, seasonality and
        individual responses to events such as receiving a positive rapid test. For
        weekly contacts, individuals are assigned to up to four groups that are time
        constant and that meet weekly. The day on which meetings take place varies
        between groups but stays the same for each group.
        % work contacts
        The lower row shows the distribution of the different types of work contacts.
        Work contacts only take place between working individuals.
        % non recurrent
        In the model it is sampled every day which number of non recurrent work contacts
        a working person is planned to have. Note that the contact diaries include such
        high values that super spreading events are well possible in our model. The
        planned number of contacts is reduced by policies, seasonality and individual
        responses to events such as receiving a positive rapid test to the number of
        actual contacts with transmission potential.
        % daily
        Working individuals are assigned to groups that are time constant and that meet
        daily to match the given distribution of daily work contacts. This contact type
        covers contacts such as colleagues with which one shares an office space. The
        share of individuals who attend in a way that has transmission potential is
        reduced by policies (such as a work from home mandate), seasonality and
        individual responses to events such as receiving a positive rapid test.
        % weekly
        Working individuals are assigned to up to 14 groups that are time constant and
        meet weekly. Groups are scheduled to meet on separate days of the work week.
        These contact models cover weekly team meetings etc. The share of individuals
        that attend in a way that has transmission potential is reduced by policies,
        seasonality and individual responses to events such as receiving a positive rapid
        test.}
\end{figure}

An exception where we do not rely on the data by \cite{Mossong2008} are the household
contacts. Since household are included in the the German microcensus
\citep{FDSAeDBUDL2018} on which we build our synthetic population we simply assume for
the household contact model that individuals meet all other household members every day.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/household}
    \caption{Number of Household Contacts}
    \label{n_contacts_hh}
    \floatfoot{\noindent
        \textit{Note:} Every individual meets all other household members every day. The
        German microcensus sampled full households such that our synthetic population
        automatically fits population characteristics such as size and age distribution.}
\end{figure}

@@

>>>>

We calibrate the parameters for the predicted numbers of contacts from contact diaries of
over 2000 individuals from Germany, Belgium, the Netherlands and Luxembourg
\citep{Mossong2008}. Each contact diary contains all contacts an individual had
throughout one day, including information on the other person (such as age and gender)
and information on the contact. Importantly, for each contact individuals entered of
which type the contact (school, leisure, work etc.) was and how frequent the contact with
the other person is. Binning the number of contacts for very high numbers, we arrive at
the distributions of the numbers of contacts by contact type as shown in Figure~\ref{fig:n_contacts}.

\begin{figure}[ht]
    \centering
    % other contacts
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_non_recurrent}
        \caption{Number of Non Recurrent Other Contacts}
        \label{fig:n_contacts_other_non_recurrent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_daily}
        \caption{Number of Daily Recurrent Other Contacts}
        \label{fig:n_contacts_other_daily_recurrent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/other_recurrent_weekly}
        \caption{Number of Weekly Recurrent Other Contacts}
        \label{fig:n_contacts_other_weekly_recurrent}
    \end{subfigure}

@@

    % work contacts
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_non_recurrent}
        \caption{Number of Non Recurrent Work Contacts}
        \label{fig:n_contacts_work_non_recurrent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_daily}
        \caption{Number of Daily Recurrent Work Contacts}
        \label{fig:n_contacts_work_daily_recurrent}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/work_recurrent_weekly}
        \caption{Number of Weekly Recurrent Work Contacts}
        \label{fig:n_contacts_work_weekly_recurrent}
    \end{subfigure}
    \vskip3ex

    \caption{Number of Contacts of the Different Contact Types}
    \label{fig:n_contacts}
    \floatfoot{\noindent
        \textit{Note:} This figure shows the pre-pandemic number of contacts individuals
        report of different contact types.
        % non recurrent contacts
        In the model it is sampled every day which of the numbers of non recurrent
        contacts a person is planned to have. Note that the contact diaries include such
        high values that super spreading events are well possible in our model through
        non recurrent models.
        % recurrent contacts
        For recurrent contacts individuals are put into groups that meet either every day
        or on a particular week day every day. For work contacts, meetings can only take
        place on work days.
        % reductions that can happen
        The pre-pandemic number of contacts with transmission potential is reduced by
        policies, seasonality and individual responses to events such as
        receiving a positive rapid test to the number of actual contacts with
        transmission potential.
        % other
        The upper row shows the distribution of the number of other contacts individuals
        report. Other contacts include all contacts that are not household members,
        school contacts or work contacts, for example leisure contacts. We assume that
        individuals in households with children or teachers or retired individuals have
        additional non recurrent other contacts during school vacations to cover things
        like family visits or travel during vacations.
        % work contacts
        The lower row shows the distribution of the different types of work contacts.
        Work contacts only take place between working individuals.}
\end{figure}

@@

An exception where we do not rely on the data by \cite{Mossong2008} are the household
contacts. Since households are included in the the German microcensus
\citep{FDSAeDBUDL2018} on which we build our synthetic population we simply assume for
the household contacts that individuals meet all other household members every day. The
number of household contacts that happen every day is shown in
Figure~\ref{fig:n_contacts_hh}.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/results/figures/data/distributions_of_the_number_of_contacts/household}
    \caption{Number of Household Contacts}
    \label{fig:n_contacts_hh}
    \floatfoot{\noindent
        \textit{Note:} Every individual meets all other household members every day. The
        German microcensus sampled full households such that our synthetic population
        automatically fits population characteristics such as size and age distribution.}
\end{figure}

### 2021-07-11 src/paper/data/synthetic_population.tex

% workers
Workers are assigned to a daily meeting work group. The group sizes vary to match the
number of daily repeating work contacts reported by working individuals in
\cite{Mossong2008}. These groups only consist of workers that work in the same county.
For a distribution of the number of daily recurring work contacts see
Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
we match each worker with up to 14 other workers into pairs to match the number of
reported weekly work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}.
Each pair is assigned a weekday on which they always meet in the absence of work
policies. 80\% of these contacts are individuals from the same county.% work contact priority
In the same way children have an educational priority determining if they are entitled to
emergency care workers are assigned a work contact priority that captures how necessary
their work is and to which degree they can work from home. This means that it's always
the same individuals that continue to have work contacts when work from home mandates of
a certain strictness are in place. We use the reduction in work mobility reported by
\cite{Google2021} as a proxy for the share of individuals that do not have work contacts.

@@

Lastly, for the distribution of vaccinations every individual is assigned a vaccination
group and a vaccination rank from that group that creates a complete vaccination queue
over the population including 15\% that refuse to be vaccinated as reported by the Robert
Koch Institute (RKI, \cite{RKI2021}). The vaccination groups are created to match the
recommendations by the Ständige Impfkommission \citep{VygenBonnet2020}.\footnote{We cover
that teachers were prioritized more than recommended by the commission.} To cover that
the Pfizer-BioNTech vaccine was later approved for younger age groups we put adolescents
and children into two groups that follow after the general population. These groups do
not become eligible within our simulation frame until June. The way vaccinations are rolled out in our model is shown in Figure~\ref{fig:vaccinations_by_age_group}.

>>>>

% workers
Workers are assigned to a daily meeting work group. The group sizes vary to match the
number of daily repeating work contacts reported by working individuals in
\cite{Mossong2008}. These groups only consist of workers that work in the same county.
For a distribution of the number of daily recurring work contacts see
Figure~\ref{n_contacts_work_daily_recurrent}. To match the number of weekly work groups
we match each worker with up to 14 other workers into pairs to match the number of
reported weekly work contacts shown in Figure~\ref{n_contacts_work_weekly_recurrent}.
Each pair is assigned a weekday on which they always meet in the absence of work
policies. 80\% of these contacts are individuals from the same county.
% work contact priority
In the same way children have an educational priority determining if they are entitled to
emergency care workers are assigned a work contact priority that captures how necessary
their work is and to which degree they can work from home. This means that it's always
the same individuals that continue to have work contacts when work from home mandates of
a certain strictness are in place. We use the reduction in work mobility reported by
\cite{Google2021} as a proxy for the share of individuals that do not have work contacts.

@@

Lastly, for the distribution of vaccinations every individual is assigned a vaccination
group and a vaccination rank from that group that creates a complete vaccination queue
over the population including 15\% that refuse to be vaccinated as reported by the
Robert Koch Institute (RKI, \cite{RKI2021c}). The vaccination groups are created to
match the recommendations by the Ständige Impfkommission
\citep{VygenBonnet2020}.\footnote{We cover that teachers were prioritized more than
recommended by the commission.} To cover that the Pfizer-BioNTech vaccine was later
approved for younger age groups we put adolescents and children into two groups that
follow after the general population. These groups do not become eligible within our
simulation frame until June. The way vaccinations are rolled out in our model is shown
in Figure~\ref{fig:vaccinations_by_age_group}.

### 2021-07-11 src/paper/model/testing.tex

To distribute the correct number of PCR tests among symptomatic and asymptomatic
infections without knowing explicit test demand probabilities, we use the following
approach: First, we calculate the total number of positive PCR tests by multiplying the
number of newly infected individuals with an estimate of the share of detected cases from
the Dunkelzifferradar project \citep{Dunkelzifferradar2020}. Next, we determine how many of these tests
should go to symptomatic and asymptomatic individuals from data by the Robert Koch
Institute. \comment[id=J]{add source}
Then, we sample the individuals to which those tests are allocated from the pools of
symptomatic and asymptomatic infected but not yet tested individuals.

>>>>

To distribute the correct number of PCR tests among symptomatic and asymptomatic
infections without knowing explicit test demand probabilities, we use the following
approach: First, we calculate the total number of positive PCR tests by multiplying the
number of newly infected individuals with an estimate of the share of detected cases
from the Dunkelzifferradar project \citep{Dunkelzifferradar2020}. Next, we determine how
many of these tests should go to symptomatic and asymptomatic individuals from data by
the RKI.\comment[id=J]{add source} Then, we sample the individuals to which those tests
are allocated from the pools of symptomatic and asymptomatic infected but not yet tested
individuals.

### 2021-07-11 src/paper/results/fit.tex

\begin{figure}[ht]     % rapid test demand
    \centering
    \caption{Share of Individuals With Rapid Tests}
    \label{fig:share_ever_rapid_test2}
    \begin{subfigure}{.425\textwidth}
        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
        \caption{Share of the Population That Has Ever Done a Rapid Test}
    \end{subfigure}%
    \hfill
    \begin{subfigure}{.425\textwidth}
        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
        \caption{Share of the Population That Did a Rapid Test in the Last Week}
    \end{subfigure}
    \label{fig:share_rapid_test_last_week2}
    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
        have ever done a rapid test or done a rapid test within the last week in our
        simulations to the shares reported in the
        \href{https://projekte.uni-erfurt.de/cosmo2020/web/topic/wissen-verhalten/80-schnelltests/}{COVID-19
        Snapshot Monitoring Survey}. The left panel compares the share of individuals who
        have ever done a rapid test. The right panel compares the share of individuals
        who have done a rapid test within the last seven days in our simulation compared
        to the share reporting to have done at least weekly rapid tests in the last four
        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
        conservative. The overall share is below that in the study. We fit the share of
        weekly tests quite exactly. However, the study only covers adults while our share
        also includes children who are tested very regularly when attending school.}
\end{figure}

>>>>

\begin{figure}[ht]     % rapid test demand
    \centering
    \caption{Share of Individuals With Rapid Tests}
    \label{fig:share_ever_rapid_test2}
    \begin{subfigure}{.425\textwidth}
        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_ever_rapid_test}
        \caption{Share of the Population That Has Ever Done a Rapid Test}
    \end{subfigure}%
    \hfill
    \begin{subfigure}{.425\textwidth}
        \includegraphics[width=0.9 \textwidth]{figures/results/figures/scenario_comparisons/combined_fit/full_share_rapid_test_in_last_week}
        \caption{Share of the Population That Did a Rapid Test in the Last Week}
    \end{subfigure}
    \label{fig:share_rapid_test_last_week2}
    \floatfoot{\noindent \textit{Note:} The figure compares the share of individuals who
        have ever done a rapid test or done a rapid test within the last week in our
        simulations to the shares reported in \citet{Betsch2021b}. The left panel compares the share of individuals who
        have ever done a rapid test. The right panel compares the share of individuals
        who have done a rapid test within the last seven days in our simulation compared
        to the share reporting to have done at least weekly rapid tests in the last four
        weeks in the COSMO survey. Overall our calibration of rapid tests are slightly
        conservative. The overall share is below that in the study. We fit the share of
        weekly tests quite exactly. However, the study only covers adults while our share
        also includes children who are tested very regularly when attending school.}
\end{figure}

### 2021-07-11 src/paper/results/rapid_tests_statistics.tex

\begin{figure}[ht] % Share tested per day
  \centering
  \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
  \caption{Share of the Population Demanding a Rapid Test Because of Different Channels
  on a Given Day}
  \label{fig:rapid_test_demand_by_channel}
  \floatfoot{\noindent \textit{Note:}
    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
    school) as well as school pupils. After Easter the required frequency of tests is
    increased from once per week to twice per week. Work rapid tests are demanded by
    individuals that still have work contacts, i.e. do not work from home. The share of
    employers offering rapid tests increases over the time frame and the frequency of
    testing is also increased. Tests are demanded by individuals for one of three private
    reasons: having developed symptoms without access to a PCR test, having a household
    member that has tested positive or developed symptoms or having planned weekly
    meeting with friends. }
\end{figure}

>>>>

\begin{figure}[ht] % Share tested per day
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
      \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested}
      \caption{Share of the Population Doing a Rapid Test Because of Different Channels
      on a Given Day}
      \label{fig:rapid_test_demand_by_channel}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
      \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested_positive}
      \caption{Share of the Population Testing Positive Because of Different Channels
      on a Given Day}
      \label{fig:pos_rapid_tests_by_channel}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
      \includegraphics[width=\textwidth]{figures/results/figures/rapid_test_statistics/popshare_tested_negative}
      \caption{Share of the Population Testing Negative Because of Different Channels
      on a Given Day}
      \label{fig:neg_rapid_tests_by_channel}
  \end{subfigure}
  \caption{Rapid Test Shares in the Population by Channel}
  \floatfoot{\noindent \textit{Note:}
    Rapid tests in the education setting are demanded by teachers (nursery, preschool and
    school) as well as school pupils. After Easter the required frequency of tests is
    increased from once per week to twice per week. Work rapid tests are demanded by
    individuals that still have work contacts, i.e. do not work from home. The share of
    employers offering rapid tests increases over the time frame and the frequency of
    testing is also increased. Tests are demanded by individuals for one of three private
    reasons: having developed symptoms without access to a PCR test, having a household
    member that has tested positive or developed symptoms or having planned weekly
    meeting with friends. The left figure shows the share of the population demanding a
    rapid test on a given day. The middle figure shows the share of the population
    testing positive on a given day (true and false positives). The right figure shows
    the share of the population testing negative on a given day (true and false
    negatives).}
\end{figure}
